**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-05-25 submission/manuscript.Rmd
discussion tweaks
results

@@

We identified a range of Simpson diversity values between patients and between each anatomical site. As this measurement takes both number of species and abundance of species into account, it seems that there is a varied level of each component along the human colon. There was no clear pattern of diversity when comparing patients- some patients had a higher level of diversity in the proximal than the distal, some patients had the opposite. An early study of microbial diversity and membership of three human subjects found a varied level of diversity in each subject, but that subjects were overall more similar to themselves than to each other (Eckberg 2005). When ordinating the subject communities to compare the right and left sides, we found that more than half of the samples from the same patient clustered near each other (Figure 2C). However, there were a few patients whose samples were very different from each other and did not ordinate in that manner. One explanation for this ____ - human diversity, sampling locations arent perfectly matched (but so what?)

discussion of diversity results- different patterns in each person. ref previous studies of diversity along gut axis of healthy people. in this same paper ( I think) can also reference difference in NMDS... if I end up including that.

discussion of similarity results- may combine with previous paragraph here, prior studies, again underscore extreme variability of humans

Discussion of RF models, how they compare to FIT, Joe's check joes paper, how well does model do in L vs R for FIT?
when we look at models for CRC, cant predict R vs L, so create models

discuss host response papers? what does that mean.

@@

-what's next?  But really to understand what's going on in the various regions of the gut we need more information than just 16S

>>>>

this is a really good paper, good for framing intro and discussion : Numerical ecology validates a biogeographical distribution and gender-based effect on mucosa-associated bacteria along the human colon

@@

We identified a range of Simpson diversity values between patients and between each anatomical site. As this measurement takes both number of species and abundance of species into account, it seems that there is a varied level of each component along the human colon. There was no clear pattern of diversity when comparing patients- some patients had a higher level of diversity in the proximal than the distal, some patients had the opposite. An early study of microbial diversity and membership of three human subjects found a varied level of diversity in each subject, but that subjects were overall more similar to themselves than to each other (Eckberg 2005). We next used the thetaYC dissimilarity metric to compare the communities present at each anatomical location to each other. We found that while there was a wide range of similarity of each site within a person, the study participants were ultimately more similar to themselves than all of the other participants. This finding is consistent with numerous studies examining the biogeography of the colon in healthy humans (Eckberg 2005, Aguierre de Carcer, 2015).

To get around the noisiness created by a diverse set of samples, we employed machine learning algorithms to identify microbes specific to each side.

Random Forest models have been used by our group and others to increase the detection sensitivity of CRC tumors. Indeed, our group found that a classification model that used microbiome data in combination with Fecal Immunohistochemical Test (FIT) results could correctly identify both carcinoma and adenoma lesions from communities of stool at exit (Baxter et al 2016) and it performed much better than FIT alone. 
- Could try to predict L vs R tumors from exit stool? long term. maybe. probably not. value of getting data about exit stool comparisons. real issue is diagnosing right though. 
- note that we have FIT for all of these people but they're all below limit of detection (except for 2). when we look at models for CRC, cant predict R vs L, so create models for healthy

@@

-what's next?  But really to understand what's going on in the various regions of the gut we need more information than just 16S 
discuss host response papers? what does that mean. link bugs to host response on left and right? future directions or intro
